---
title: "tourist_flow_prediction"
author: "cuong nguyen"
date: "2023-10-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages('forecast')
install.packages('tsutils'); library(tsutils)
```

```{r}
pckg <- c("thief","MAPA","tsutils","abind") 
for (i in 1:length(pckg)){
  if(!(pckg[i] %in% rownames(installed.packages()))){ 
    install.packages(pckg[i])
  }
  library(pckg[i],character.only = TRUE)
}
```
```{r}
install.packages("dplyr")
```
```{r}
if (!require("glmnet")){install.packages("glmnet")}; library(glmnet)
```

```{r}
library(forecast)
library(tsutils)
library(dplyr)
```

```{r}
load("dataTour.Rdata")
load("IMFdata.Rdata")
```

```{r}
#create a subset from macroeconomic dataset
russiasub <- X$Russia
russiasub <- window(russiasub,start=c(1995),end=c(2019))
```

```{r}
#load the primary dataset
df <- dataTour$Finland 
#drop the year 2020 observations
df <- head(df, -1)
```
We start exploring each country time series and build accordance models 
```{r}
russia <- df[, "Russia"]
china <- df[,"China"]
germany <- df[, "Germany"]
uk <- df[, "United Kingdom"]
sweden <- df[, "Sweden"]
```

```{r}
# Create a grid of plots
par(mfrow=c(2,3))
plot(russia)
plot(china)
plot(germany)
plot(uk)
plot(sweden)
```
```{r}
par(mfrow=c(2,3))
pacf(russia)
pacf(china)
pacf(germany)
pacf(uk)
pacf(sweden)
```
```{r}
par(mfrow=c(3,1))
plot(cmav(germany),main="Germany")
plot(cmav(uk),main="UK")
plot(cmav(sweden),main="Sweden")
```


##Russia
```{r}
#split training and test sets
russia.tst <- tail(russia,4)
russia.trn <- head(russia,24*4)
#observe potential trend, seasonality
cma.russia <- cmav(russia.trn,outplot=1) #perform central moving average  
seasplot(russia.trn)  #observe seasonality
dc.russia <- decomp(russia.trn,outplot=1) #decomposition
pacf(russia.trn)
```
There is presence of trend and seasonality in the time series
```{r}
fit1 <- ets(russia.trn) 
fit1
```
We build the ETS model
```{r}
#split train set to smaller training set and validation set
russia.ins <- head(russia.trn,18*4)
russia.val <- tail(russia.trn,5*4)
#observe maximum step of prediction can be included in the validation set
h <- 4
omax <- length(russia.val) - h + 1
models <- c("AAA","ANA","MAM","MMM","Naive") 
damped <- c(TRUE, FALSE, TRUE, TRUE)
err <- array(NA,c(omax,5)) 
frcs <- array(NA,c(h,5))
#Perform rolling origin on in-sample set
for (o in 1:omax){
  russia.ins <- head(russia.trn,18*4-1+o) 
  russia.val <- tail(russia.trn,5*4-o+1) 
  for (m in 1:4){
    fit1 <- ets(russia.ins,model=models[m],damped=damped[m]) 
    frcs[,m] <- forecast(fit1,h=h)$mean
    err[o,m] <- mean(abs(russia.val[1:h] - frcs[,m]))
  }
  frcs[,5] <- tail(russia.ins,frequency(russia.ins))[1:h] 
  err[o,5] <- mean(abs(russia.val[1:h] - frcs[,5]))
}
colnames(err) <- c("AAA","ANA","MAM","MMM","Naive")
errMean <- colMeans(err)
errMean
```
```{r}
# out-of-sample validation
modelsTest <- c("AAA","ANA","MAM","MMM","MNM", "Naive", "CombMean", "CombMedian")
dampedTest <- c(TRUE, FALSE, TRUE, TRUE,FALSE)
omaxTest <- length(russia.tst) - h + 1
errTest <- array(NA,c(omaxTest,8))
frcsTest <- array(NA,c(h,8))
for (o in 1:omaxTest){
  russia.trnTest <- head(russia,21*4-1+o) 
  russia.tstTest <- tail(russia,2*4-o+1) 
    for (m in 1:5){
      fit1 <- ets(russia.trnTest,model=modelsTest[m],damped=dampedTest[m]) 
      frcsTest[,m] <- forecast(fit1,h=h)$mean
      errTest[o,m] <- mean(abs(russia.tstTest[1:h] - frcsTest[,m]))
    }
  frcsTest[,6] <- tail(russia.trnTest,frequency(russia.trnTest))[1:h]
  errTest[o,6] <- mean(abs(russia.tstTest[1:h] - frcsTest[,6]))
  frcsTest[,7] <- apply(frcsTest[,1:5],1,mean)
  errTest[o,7] <- mean(abs(russia.tstTest[1:h] - frcsTest[,7]))
  frcsTest[,8] <- apply(frcsTest[,1:5],1,median)
  errTest[o,8] <- mean(abs(russia.tstTest[1:h] - frcsTest[,8]))
}
colnames(errTest) <- c("AAA","ANA","MAM","MMM","MNM","Naive","Comb.Mean","Comb.Median")
errMeanTest <- colMeans(errTest)
errMeanTest
```

Naive gives the smallest error hence it is preferable
```{r}
mape1 <- MAPE(russia.tst,tail(russia.trnTest,frequency(russia.trnTest))[1:4]) * 100
cat("MAPE:", mape1, "%\n")
```
```{r}
#Building model with AAA
fit1 <- ets(russia.trn,model="ANA",damped=FALSE) 
frc1 <- forecast(fit1,h=4)
mape <- MAPE(russia.tst,frc1$mean)*100
cat("MAPE:", mape, "%\n")
```

Now we construct lag for regression modeling

```{r}
n <- length(russia.trn)
X <- array(NA,c(n,6))
for (i in 1:6){
  X[i:n,i] <- russia.trn[1:(n-i+1)]
}
colnames(X) <- c("y",paste0("lag",1:5))
X <- as.data.frame(X)
plot(X)
# The complete model
fit2 <- lm(y~.,data=X)
# The stepwise model
fit2 <- step(fit2)
# In-sample fit:
plot(X$y,type="l")
frc2 <- predict(fit2,X)
lines(frc2,col="red")
frc2 <- array(NA,c(4,1))
for (i in 1:4){
  Xnew <- tail(russia.trn,5)
  Xnew <- c(Xnew,frc2)
  Xnew <- Xnew[i:(4+i)]
  Xnew <- Xnew[5:1]
  Xnew <- array(Xnew, c(1,5)) 
  colnames(Xnew) <- paste0("lag",1:5)
  Xnew <- as.data.frame(Xnew)
  frc2[i] <- predict(fit2,Xnew)
} 
mape <- MAPE(russia.tst,frc2)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#dummies variables
D <- factor(rep(1:4,24))
X2 <- cbind(X,D)
colnames(X2) <- c(colnames(X2)[1:6],"D")
fit3 <- lm(y~.,data=X2)
summary(fit3)
# Find NA in X2
idx <- is.na(X2)
idx <- rowSums(idx)
idx <- idx == 0
fit_temp <- lm(y~.,data=X2[idx,])
# fit_temp is the same as fit3, without the first NA part 
fit4 <- step(fit_temp)
summary(fit4)
frc3 <- predict(fit4,X2)
ts.plot(russia.trn,frc3,col=c("black","red"))
frc3 <- array(NA,c(4,1)) 
for (i in 1:4){
  Xnew <- tail(russia.trn,5)
  Xnew <- c(Xnew,frc3)
  Xnew <- Xnew[i:(4+i)]
  Xnew <- Xnew[5:1]
  Xnew <- array(Xnew, c(1,5))
  colnames(Xnew) <- paste0("lag",1:5)
  Xnew <- as.data.frame(Xnew)
  D <- as.factor(rep(1:4,1)[i])
  frc3[i] <- predict(fit4,Xnew)
}
mape <- MAPE(russia.tst,frc3)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Lasso
xx <- as.matrix(X[-(1:6),-1])
# For the target I retain only the first column 
yy <- as.matrix(X[-(1:6),1])
lasso <- cv.glmnet(x=xx,y=yy)
frc4 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(russia.trn,5),frc4)
  Xnew <- (Xnew[i:(4+i)])[5:1]
  Xnew <- array(Xnew, c(1,5)) 
  colnames(Xnew) <- paste0("lag",1:5) 
  # Forecast
  frc4[i] <- predict(lasso,Xnew)
}
mape <- MAPE(russia.tst,frc4)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Ridge
ridge <- cv.glmnet(x=xx,y=yy,alpha=0)
frc5 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(russia.trn,5),frc5)
  Xnew <- (Xnew[i:(4+i)])[5:1]
  Xnew <- array(Xnew, c(1,5))
  colnames(Xnew) <- paste0("lag",1:5) # Forecast 
  frc5[i] <- predict(ridge,Xnew)
}
mape <- MAPE(russia.tst,frc5)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Building THieF
frc6 <- thief(russia.trn,h=4,usemodel="arima")
frc6_values <- frc6$mean
mape <- MAPE(russia.tst,frc6_values)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#building differencing model to handling trend
X3<-X
for (i in 1:ncol(X3)){
  X3[,i] <- c(NA,diff(X3[,i]))
} 
summary(lm(y~.,X3))
if (any(is.na(X3))) {
  # Handle missing values 
  dataset_clean <- na.omit(X3)  
  # Re-run analysis, including stepwise model selection
  fit5 <- lm(y~., data = dataset_clean)
  fit6 <- step(fit5)
} else {
  # No missing values, so proceed with analysis
  fit5 <- lm(y~., data = X3)
  fit6 <- step(fit5)
}
frc7 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Calculate the differences of the in-sample data
  russia.diff <- diff(russia.trn)
  # Create lags - same as before
  Xnew <- tail(russia.diff,5)
  Xnew <- c(Xnew,frc7)
  Xnew <- Xnew[i:(4+i)]
  Xnew <- Xnew[5:1]
  Xnew <- array(Xnew, c(1,5)) 
  colnames(Xnew) <- paste0("lag",1:5) 
  Xnew <- as.data.frame(Xnew)
  # Forecast
  frc7[i] <- predict(fit6,Xnew)
}
```

```{r}
# Transform to time series
frc7 <- ts(frc7,frequency=frequency(russia.tst),start=start(russia.tst)) # Plot
ts.plot(diff(russia.trn),frc7,col=c("black","red"))
```

```{r}
frc7ud <- cumsum(c(tail(russia.trn,1),frc7))
frc7ud <- frc7ud[-1]
```

```{r}
frc2 <- ts(frc2,frequency=frequency(russia.tst),start=start(russia.tst))
frc3 <- ts(frc3,frequency=frequency(russia.tst),start=start(russia.tst))
frc4 <- ts(frc4,frequency=frequency(russia.tst),start=start(russia.tst))
frc5 <- ts(frc5,frequency=frequency(russia.tst),start=start(russia.tst))
frc6 <- ts(frc6_values,frequency=frequency(russia.tst),start=start(russia.tst))
frc7ud <- ts(frc7ud,frequency=frequency(russia.tst),start=start(russia.tst))

# Plot
ts.plot(tail(russia.trn,16), russia.tst, frc1$mean, frc2, frc3, frc4, frc5, frc6, frc7ud, col = c("black", "black", "red", "blue","green","yellow","purple","orange","brown"))
legend("bottomleft", c("ETS(ANA)","Stepwise", "Dummies", "Lasso", "Ridge","Thief", "Differencing"), col = c("red", "blue","green","yellow","purple","orange","brown"), lty = 1)
```
China
```{r}
#split training and test sets
china.tst <- tail(china,4)
china.trn <- head(china,4*24)
#observe potential trend, seasonality
cma.china <- cmav(china.trn,outplot=1) #seems like the ts trend 
seasplot(china.trn) #there is no sign of seasonality 
dc.china <- decomp(china.trn,outplot=1) #the plot shows that there is appearance of seasonality
pacf(china.trn)
```
```{r}
fit1 <- ets(china.trn) 
fit1
```

```{r}
#Building ETS on training set
#split train set to smaller training set and validation set
china.ins <- head(china.trn,18*4)
china.val <- tail(china.trn,5*4)
h <- 4
#observe maximum step of prediction can be included in the validation set
omax <- length(china.val) - h + 1
models <- c("AAA","ANA","MMM","MAM","Naive") 
damped <- c(TRUE, FALSE, TRUE,TRUE)
err <- array(NA,c(omax,5)) 
frcs <- array(NA,c(h,5))
#Perform rolling origin on in-sample set
for (o in 1:omax){
  china.ins <- head(china.trn,19*4-1+o) 
  china.val <- tail(china.trn,5*4-o+1) 
  for (m in 1:4){
    fit1 <- ets(china.ins,model=models[m],damped=damped[m]) 
    frcs[,m] <- forecast(fit1,h=h)$mean
    err[o,m] <- mean(abs(china.val[1:h] - frcs[,m]))/100
  }
  frcs[,5] <- tail(china.ins,frequency(china.ins))[1:h] 
  err[o,5] <- mean(abs(china.val[1:h] - frcs[,5]))/100
}
colnames(err) <- c("AAA","ANA","MMM","MAM","Naive")
errMean <- colMeans(err)
errMean
```

```{r}
# Building ETS on test set
modelsTest <- c("AAA","ANA","MAM","MMM", "Naive", "CombMean", "CombMedian")
dampedTest <- c(TRUE, FALSE, TRUE,TRUE)
omaxTest <- length(china.tst) - h + 1
errTest <- array(NA,c(omaxTest,7))
frcsTest <- array(NA,c(h,7))
for (o in 1:omaxTest){
  china.trnTest <- head(china,23*4-1+o) 
  china.tstTest <- tail(china,2*4-o+1) 
    for (m in 1:4){
      fit1 <- ets(china.trnTest,model=modelsTest[m],damped=dampedTest[m]) 
      frcsTest[,m] <- forecast(fit1,h=h)$mean
      errTest[o,m] <- mean(abs(china.tstTest[1:h] - frcsTest[,m]))/100
    }
  frcsTest[,5] <- tail(china.trnTest,frequency(china.trnTest))[1:h]
  errTest[o,5] <- mean(abs(china.tstTest[1:h] - frcsTest[,5]))/100
  frcsTest[,6] <- apply(frcsTest[,1:5],1,mean)
  errTest[o,6] <- mean(abs(china.tstTest[1:h] - frcsTest[,6]))/100
  frcsTest[,7] <- apply(frcsTest[,1:5],1,median)
  errTest[o,7] <- mean(abs(china.tstTest[1:h] - frcsTest[,7]))/100
}
colnames(errTest) <- c("AAA","ANA","MAM","MMMd","Naive","Comb.Mean","Comb.Median")
errMeanTest <- colMeans(errTest)
errMeanTest
```

AAA model performs the best in both training and test set

```{r}
#Building model with AAA
fit1 <- ets(china.trn,model="AAA",damped=TRUE) 
frc1 <- forecast(fit1,h=4)
mape <- MAPE(china.tst,frc1$mean)*100
cat("MAPE:", mape, "%\n")
```

Construct lag 
```{r}
n <- length(china.trn)
X <- array(NA,c(n,7))
for (i in 1:7){
  X[i:n,i] <- china.trn[1:(n-i+1)]
}
colnames(X) <- c("y",paste0("lag",1:6))
X <- as.data.frame(X)
plot(X)
# The complete model
fit2 <- lm(y~.,data=X)
# The stepwise model
fit2 <- step(fit2)
# In-sample fit:
plot(X$y,type="l")
frc2 <- predict(fit2,X)
lines(frc2,col="red")
frc2 <- array(NA,c(4,1))
for (i in 1:4){
  Xnew <- tail(china.trn,6)
  Xnew <- c(Xnew,frc2)
  Xnew <- Xnew[i:(5+i)]
  Xnew <- Xnew[6:1]
  Xnew <- array(Xnew, c(1,6)) 
  colnames(Xnew) <- paste0("lag",1:6)
  Xnew <- as.data.frame(Xnew)
  frc2[i] <- predict(fit2,Xnew)
} 
mape <- MAPE(china.tst,frc2)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Dummies variable
D <- factor(rep(1:4,24))
X2 <- cbind(X,D)
colnames(X2) <- c(colnames(X2)[1:7],"D")
fit3 <- lm(y~.,data=X2)
summary(fit3)
# Find NA in X2
idx <- is.na(X2)
idx <- rowSums(idx)
idx <- idx == 0
fit_temp <- lm(y~.,data=X2[idx,])
# fit_temp is the same as fit3, without the first NA part 
fit4 <- step(fit_temp)
summary(fit4)
frc3 <- array(NA,c(4,1)) 
for (i in 1:4){
  Xnew <- tail(china.trn,6)
  Xnew <- c(Xnew,frc3)
  Xnew <- Xnew[i:(5+i)]
  Xnew <- Xnew[6:1]
  Xnew <- array(Xnew, c(1,6))
  colnames(Xnew) <- paste0("lag",1:6)
  Xnew <- as.data.frame(Xnew)
  D <- as.factor(rep(1:4,1)[i])
  frc3[i] <- predict(fit4,Xnew)
}
mape <- MAPE(china.tst,frc3)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Lasso
xx <- as.matrix(X[-(1:6),-1])
# For the target I retain only the first column 
yy <- as.matrix(X[-(1:6),1])
lasso <- cv.glmnet(x=xx,y=yy)
frc4 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(china.trn,6),frc4)
  Xnew <- (Xnew[i:(5+i)])[6:1]
  Xnew <- array(Xnew, c(1,6)) 
  colnames(Xnew) <- paste0("lag",1:6) 
  # Forecast
  frc4[i] <- predict(lasso,Xnew)
}
mape <- MAPE(china.tst,frc4)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Ridge
ridge <- cv.glmnet(x=xx,y=yy,alpha=0)
frc5 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(china.trn,6),frc5)
  Xnew <- (Xnew[i:(5+i)])[6:1]
  Xnew <- array(Xnew, c(1,6))
  colnames(Xnew) <- paste0("lag",1:6) # Forecast 
  frc5[i] <- predict(ridge,Xnew)
}
mape <- MAPE(china.tst,frc5)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Building THieF
frc6 <- thief(china.trn,h=4,usemodel="arima")
frc6_values <- frc6$mean
mape <- MAPE(china.tst,frc6_values)*100
cat("MAPE:", mape, "%\n")
```

```{r}
frc2 <- ts(frc2,frequency=frequency(china.tst),start=start(china.tst))
frc3 <- ts(frc3,frequency=frequency(china.tst),start=start(china.tst))
frc4 <- ts(frc4,frequency=frequency(china.tst),start=start(china.tst))
frc5 <- ts(frc5,frequency=frequency(china.tst),start=start(china.tst))
frc6 <- ts(frc6_values,frequency=frequency(china.tst),start=start(china.tst))
# Plot
ts.plot(tail(china.trn,4), china.tst, frc1$mean, frc2, frc3, frc4, frc5, frc6, col = c("black", "black", "red", "blue","green","purple","orange","brown"))
legend("bottomleft", c("ETS(AAA)","Stepwise", "Dummies","Lasso","Ridge","Thief"), col = c("red", "blue","green","purple","orange","brown"), lty = 1)
```

Examine germany time series
```{r}
#split training and test sets
germany.tst <- tail(germany,4)
germany.trn <- head(germany,4*24)
#observe potential trend, seasonality
cma.germany <- cmav(germany.trn,outplot=1) #seems like the ts trend 
seasplot(germany.trn) #there is no sign of seasonality 
dc.germany <- decomp(germany.trn,outplot=1) #the plot shows that there is appearance of seasonality
pacf(germany.trn)
```
```{r}
fit1 <- ets(germany.trn) 
fit1
```

```{r}
#Building ETS on training set
#split train set to smaller training set and validation set
germany.ins <- head(germany.trn,18*4)
germany.val <- tail(germany.trn,5*4)
h <- 4
#observe maximum step of prediction can be included in the validation set
omax <- length(germany.val) - h + 1
models <- c("AAA","ANA","MNM","MAM","Naive") 
damped <- c(TRUE, FALSE, FALSE,TRUE)
err <- array(NA,c(omax,5)) 
frcs <- array(NA,c(h,5))
#Perform rolling origin on in-sample set
for (o in 1:omax){
  germany.ins <- head(germany.trn,19*4-1+o) 
  germany.val <- tail(germany.trn,5*4-o+1) 
  for (m in 1:4){
    fit1 <- ets(germany.ins,model=models[m],damped=damped[m]) 
    frcs[,m] <- forecast(fit1,h=h)$mean
    err[o,m] <- mean(abs(germany.val[1:h] - frcs[,m]))
  }
  frcs[,5] <- tail(germany.ins,frequency(germany.ins))[1:h] 
  err[o,5] <- mean(abs(germany.val[1:h] - frcs[,5]))
}
colnames(err) <- c("AAA","ANA","MNM","MAM","Naive")
errMean <- colMeans(err)
errMean
```
```{r}
# Building ETS on test set
modelsTest <- c("AAA","ANA","MNM","MAM", "Naive", "CombMean", "CombMedian")
dampedTest <- c(TRUE, FALSE, FALSE,TRUE)
omaxTest <- length(germany.tst) - h + 1
errTest <- array(NA,c(omaxTest,7))
frcsTest <- array(NA,c(h,7))
for (o in 1:omaxTest){
  germany.trnTest <- head(germany,23*4-1+o) 
  germany.tstTest <- tail(germany,2*4-o+1) 
    for (m in 1:4){
      fit1 <- ets(germany.trnTest,model=modelsTest[m],damped=dampedTest[m]) 
      frcsTest[,m] <- forecast(fit1,h=h)$mean
      errTest[o,m] <- mean(abs(germany.tstTest[1:h] - frcsTest[,m]))/100
    }
  frcsTest[,5] <- tail(germany.trnTest,frequency(germany.trnTest))[1:h]
  errTest[o,5] <- mean(abs(germany.tstTest[1:h] - frcsTest[,5]))
  frcsTest[,6] <- apply(frcsTest[,1:5],1,mean)
  errTest[o,6] <- mean(abs(germany.tstTest[1:h] - frcsTest[,6]))
  frcsTest[,7] <- apply(frcsTest[,1:5],1,median)
  errTest[o,7] <- mean(abs(germany.tstTest[1:h] - frcsTest[,7]))
}
colnames(errTest) <- c("AAA","ANA","MNM","MAM","Naive","Comb.Mean","Comb.Median")
errMeanTest <- colMeans(errTest)
errMeanTest
```
```{r}
#Building model with MNM
fit1 <- ets(germany.trn,model="MNM",damped=FALSE) 
frc1 <- forecast(fit1,h=4)
mape <- MAPE(germany.tst,frc1$mean)*100
cat("MAPE:", mape, "%\n")
```
```{r}
n <- length(germany.trn)
X <- array(NA,c(n,5))
for (i in 1:5){
  X[i:n,i] <- germany.trn[1:(n-i+1)]
}
colnames(X) <- c("y",paste0("lag",1:4))
X <- as.data.frame(X)
plot(X)
```
```{r}
fit2 <- lm(y~.,data=X)
fit2 <- step(fit2)
#Forecast on training set
plot(X$y,type="l")
frc2 <- predict(fit2,X)
lines(frc2,col="red")
#Forecast on testing set
frc2 <- array(NA,c(4,1))
for (i in 1:4){
  # For the Xnew we use the last five observations as before
  Xnew <- tail(germany.trn,4)
  # Add to that the forecasted values
  Xnew <- c(Xnew,frc2)
  # Take the relevant 5 values. The index i helps us to get the right ones 
  Xnew <- Xnew[i:(3+i)]
  # If i = 1 then this becomes Xnew[1:5].
  # If i = 2 then this becomes Xnew[2:6] - just as the example above.
  # Reverse the order
  Xnew <- Xnew[4:1]
  # Make Xnew an array and name the inputs
  Xnew <- array(Xnew, c(1,4)) # c(1,5) are the dimensions of the array 
  colnames(Xnew) <- paste0("lag",1:4) # I have already reversed the order 
  # Convert to data.frame
  Xnew <- as.data.frame(Xnew)
  # Forecast
  frc2[i] <- predict(fit2,Xnew)
} 
mape <- MAPE(germany.tst,frc2)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Building dummies variable model
D <- factor(rep(1:4,24))
X2 <- cbind(X,D)
colnames(X2) <- c(colnames(X2)[1:5],"D")
fit3 <- lm(y~.,data=X2)
summary(fit3)
# Find NA in X2
idx <- is.na(X2)
idx <- rowSums(idx)
idx <- idx == 0
fit_temp <- lm(y~.,data=X2[idx,])
# fit_temp is the same as fit3, without the first NA part 
fit4 <- step(fit_temp)
summary(fit4)
frc3 <- array(NA,c(4,1)) 
for (i in 1:4){
  Xnew <- tail(germany.trn,4)
  Xnew <- c(Xnew,frc3)
  Xnew <- Xnew[i:(3+i)]
  Xnew <- Xnew[4:1]
  Xnew <- array(Xnew, c(1,4))
  colnames(Xnew) <- paste0("lag",1:4)
  Xnew <- as.data.frame(Xnew)
  D <- as.factor(rep(1:4,1)[i])
  frc3[i] <- predict(fit4,Xnew)
}
mape <- MAPE(germany.tst,frc3)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Lasso
xx <- as.matrix(X[-(1:4),-1])
# For the target I retain only the first column 
yy <- as.matrix(X[-(1:4),1])
lasso <- cv.glmnet(x=xx,y=yy)
frc4 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(germany.trn,4),frc4)
  Xnew <- (Xnew[i:(3+i)])[4:1]
  Xnew <- array(Xnew, c(1,4)) 
  colnames(Xnew) <- paste0("lag",1:4) 
  # Forecast
  frc4[i] <- predict(lasso,Xnew)
}
mape <- MAPE(germany.tst,frc4)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Ridge
ridge <- cv.glmnet(x=xx,y=yy,alpha=0)
frc5 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(russia.trn,4),frc5)
  Xnew <- (Xnew[i:(4+i)])[4:1]
  Xnew <- array(Xnew, c(1,4))
  colnames(Xnew) <- paste0("lag",1:4) # Forecast 
  frc5[i] <- predict(ridge,Xnew)
}
mape <- MAPE(germany.tst,frc5)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Building THieF
frc6 <- thief(germany.trn,h=4,usemodel="arima")
frc6_values <- frc6$mean
mape <- MAPE(germany.tst,frc6_values)*100
cat("MAPE:", mape, "%\n")
```
```{r}
frc2 <- ts(frc2,frequency=frequency(germany.tst),start=start(germany.tst))
frc3 <- ts(frc3,frequency=frequency(germany.tst),start=start(germany.tst))
frc4 <- ts(frc4,frequency=frequency(germany.tst),start=start(germany.tst))
frc5 <- ts(frc5,frequency=frequency(germany.tst),start=start(germany.tst))
frc6 <- ts(frc6_values,frequency=frequency(germany.tst),start=start(germany.tst))
# Plot
ts.plot(tail(germany.trn,4), germany.tst, frc1$mean, frc2, frc3, frc4, frc5, frc6, col = c("black", "black", "red", "blue","green","purple", "orange", "brown"))
legend("bottomleft", c("ETS(MNM)","Stepwise", "Dummies", "Lasso", "Ridge","Thief"), col = c("red", "blue","green", "purple", "orange", "brown"), lty = 1)
```

Examine UK time series
```{r}
#split training and test sets
uk.tst <- tail(uk,4)
uk.trn <- head(uk,4*24)
#observe potential trend, seasonality
cma.uk <- cmav(uk.trn,outplot=1) #seems like the ts trend 
seasplot(uk.trn) #there is no sign of seasonality 
dc.uk <- decomp(uk.trn,outplot=1) #the plot shows that there is appearance of seasonality
pacf(uk.trn)
```
```{r}
fit1 <- ets(uk.trn) 
fit1
```
```{r}
#Building ETS on training set
#split train set to smaller training set and validation set
uk.ins <- head(uk.trn,18*4)
uk.val <- tail(uk.trn,5*4)
h <- 4
#observe maximum step of prediction can be included in the validation set
omax <- length(uk.val) - h + 1
models <- c("AAA","ANA","MNM","MAM","Naive") 
damped <- c(TRUE, FALSE, FALSE,FALSE)
err <- array(NA,c(omax,5)) 
frcs <- array(NA,c(h,5))
#Perform rolling origin on in-sample set
for (o in 1:omax){
  uk.ins <- head(uk.trn,19*4-1+o) 
  uk.val <- tail(uk.trn,5*4-o+1) 
  for (m in 1:4){
    fit1 <- ets(uk.ins,model=models[m],damped=damped[m]) 
    frcs[,m] <- forecast(fit1,h=h)$mean
    err[o,m] <- mean(abs(uk.val[1:h] - frcs[,m]))
  }
  frcs[,5] <- tail(uk.ins,frequency(uk.ins))[1:h] 
  err[o,5] <- mean(abs(uk.val[1:h] - frcs[,5]))
}
colnames(err) <- c("AAA","ANA","MNM","MAM","Naive")
errMean <- colMeans(err)
errMean
```
```{r}
# Building ETS on test set
modelsTest <- c("AAA","ANA","MNM","MAM", "Naive", "CombMean", "CombMedian")
dampedTest <- c(TRUE, FALSE, FALSE,FALSE)
omaxTest <- length(uk.tst) - h + 1
errTest <- array(NA,c(omaxTest,7))
frcsTest <- array(NA,c(h,7))
for (o in 1:omaxTest){
  uk.trnTest <- head(uk,23*4-1+o) 
  uk.tstTest <- tail(uk,2*4-o+1) 
    for (m in 1:4){
      fit1 <- ets(uk.trnTest,model=modelsTest[m],damped=dampedTest[m]) 
      frcsTest[,m] <- forecast(fit1,h=h)$mean
      errTest[o,m] <- mean(abs(uk.tstTest[1:h] - frcsTest[,m]))
    }
  frcsTest[,5] <- tail(uk.trnTest,frequency(uk.trnTest))[1:h]
  errTest[o,5] <- mean(abs(uk.tstTest[1:h] - frcsTest[,5]))
  frcsTest[,6] <- apply(frcsTest[,1:5],1,mean)
  errTest[o,6] <- mean(abs(uk.tstTest[1:h] - frcsTest[,6]))
  frcsTest[,7] <- apply(frcsTest[,1:5],1,median)
  errTest[o,7] <- mean(abs(uk.tstTest[1:h] - frcsTest[,7]))
}
colnames(errTest) <- c("AAA","ANA","MNM","MAM","Naive","Comb.Mean","Comb.Median")
errMeanTest <- colMeans(errTest)
errMeanTest
```
```{r}
#Building model with MAM
fit1 <- ets(uk.trn,model="MAM",damped=FALSE) 
frc1 <- forecast(fit1,h=4)
mape <- MAPE(uk.tst,frc1$mean)*100
cat("MAPE:", mape, "%\n")
```
```{r}
n <- length(uk.trn)
X <- array(NA,c(n,8))
for (i in 1:8){
  X[i:n,i] <- uk.trn[1:(n-i+1)]
}
colnames(X) <- c("y",paste0("lag",1:7))
X <- as.data.frame(X)
plot(X)
if (any(is.na(X))) {
  # Handle missing values 
  dataset_clean <- na.omit(X)  
  # Re-run analysis, including stepwise model selection
  fit2 <- lm(y~.,data=dataset_clean)
  fit2 <- step(fit2)
} else {
  # No missing values, so proceed with analysis
  fit2 <- lm(y~.,data=dataset_clean)
  fit2 <- step(fit2)
}
#Forecast on training set
plot(X$y,type="l")
frc2 <- predict(fit2,X)
lines(frc2,col="red")
#Forecast on testing set
frc2 <- array(NA,c(4,1))
for (i in 1:4){
  Xnew <- tail(uk.trn,7)
  Xnew <- c(Xnew,frc2)
  Xnew <- Xnew[i:(6+i)]
  Xnew <- Xnew[7:1]
  Xnew <- array(Xnew, c(1,7)) 
  colnames(Xnew) <- paste0("lag",1:7)  
  Xnew <- as.data.frame(Xnew)
  frc2[i] <- predict(fit2,Xnew)
} 
mape <- MAPE(uk.tst,frc2)*100
cat("MAPE:", mape, "%\n")
```
```{r}
D <- factor(rep(1:4,24))
X2 <- cbind(X,D)
colnames(X2) <- c(colnames(X2)[1:8],"D")
fit3 <- lm(y~.,data=X2)
summary(fit3)
# Find NA in X2
idx <- is.na(X2)
idx <- rowSums(idx)
idx <- idx == 0
fit_temp <- lm(y~.,data=X2[idx,])
# fit_temp is the same as fit3, without the first NA part 
fit4 <- step(fit_temp)
summary(fit4)
frc3 <- array(NA,c(4,1)) 
for (i in 1:4){
  Xnew <- tail(uk.trn,7)
  Xnew <- c(Xnew,frc3)
  Xnew <- Xnew[i:(6+i)]
  Xnew <- Xnew[7:1]
  Xnew <- array(Xnew, c(1,7))
  colnames(Xnew) <- paste0("lag",1:7)
  Xnew <- as.data.frame(Xnew)
  D <- as.factor(rep(1:4,1)[i])
  frc3[i] <- predict(fit4,Xnew)
}
mape <- MAPE(uk.tst,frc3)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Lasso
xx <- as.matrix(X[-(1:7),-1])
# For the target I retain only the first column 
yy <- as.matrix(X[-(1:7),1])
lasso <- cv.glmnet(x=xx,y=yy)
frc4 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(uk.trn,7),frc4)
  Xnew <- (Xnew[i:(6+i)])[7:1]
  Xnew <- array(Xnew, c(1,7)) 
  colnames(Xnew) <- paste0("lag",1:7) 
  # Forecast
  frc4[i] <- predict(lasso,Xnew)
}
mape <- MAPE(uk.tst,frc4)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Ridge
ridge <- cv.glmnet(x=xx,y=yy,alpha=0)
frc5 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(uk.trn,7),frc5)
  Xnew <- (Xnew[i:(6+i)])[7:1]
  Xnew <- array(Xnew, c(1,7))
  colnames(Xnew) <- paste0("lag",1:7) # Forecast 
  frc5[i] <- predict(ridge,Xnew)
}
mape <- MAPE(uk.tst,frc5)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Building THieF
frc6 <- thief(uk.trn,h=4,usemodel="arima")
frc6_values <- frc6$mean
mape <- MAPE(uk.tst,frc6_values)*100
cat("MAPE:", mape, "%\n")
```

```{r}
frc2 <- ts(frc2,frequency=frequency(uk.tst),start=start(uk.tst))
frc3 <- ts(frc3,frequency=frequency(uk.tst),start=start(uk.tst))
frc4 <- ts(frc4,frequency=frequency(uk.tst),start=start(uk.tst))
frc5 <- ts(frc5,frequency=frequency(uk.tst),start=start(uk.tst))
frc6 <- ts(frc6_values,frequency=frequency(uk.tst),start=start(uk.tst))
# Plot
ts.plot(tail(uk.trn,4), uk.tst, frc1$mean, frc2, frc3, frc4, frc5, frc6, col = c("black", "black", "red", "blue","green", "purple", "orange","brown"))
legend("bottomleft", c("ETS(MAM)","Stepwise", "Dummies", "Lasso", "Ridge","Thief"), col = c("red", "blue","green", "purple", "orange","brown"), lty = 1)
```

```{r}
#split training and test sets
sweden.tst <- tail(sweden,4)
sweden.trn <- head(sweden,4*24)
#observe potential trend, seasonality
cma.sweden <- cmav(sweden.trn,outplot=1) #seems like the ts trend 
seasplot(sweden.trn) #there is no sign of seasonality 
dc.sweden <- decomp(sweden.trn,outplot=1) #the plot shows that there is appearance of seasonality
pacf(sweden.trn)
```
```{r}
fit1 <- ets(sweden.trn) 
fit1
```
```{r}
#Building ETS on training set
#split train set to smaller training set and validation set
sweden.ins <- head(sweden.trn,19*4)
sweden.val <- tail(sweden.trn,5*4)
h <- 4
#observe maximum step of prediction can be included in the validation set
omax <- length(sweden.val) - h + 1
models <- c("AAA","ANA","MNM","MAM","Naive") 
damped <- c(TRUE, FALSE, FALSE,FALSE)
err <- array(NA,c(omax,5)) 
frcs <- array(NA,c(h,5))
#Perform rolling origin on in-sample set
for (o in 1:omax){
  sweden.ins <- head(sweden.trn,19*4-1+o) 
  sweden.val <- tail(sweden.trn,5*4-o+1) 
  for (m in 1:4){
    fit1 <- ets(sweden.ins,model=models[m],damped=damped[m]) 
    frcs[,m] <- forecast(fit1,h=h)$mean
    err[o,m] <- mean(abs(sweden.val[1:h] - frcs[,m]))
  }
  frcs[,5] <- tail(sweden.ins,frequency(sweden.ins))[1:h] 
  err[o,5] <- mean(abs(sweden.val[1:h] - frcs[,5]))
}
colnames(err) <- c("AAA","ANA","MNM","MAM","Naive")
errMean <- colMeans(err)
errMean
```
```{r}
# Building ETS on test set
modelsTest <- c("AAA","ANA","MNM","MAM", "Naive", "CombMean", "CombMedian")
dampedTest <- c(TRUE, FALSE, FALSE,FALSE)
omaxTest <- length(sweden.tst) - h + 1
errTest <- array(NA,c(omaxTest,7))
frcsTest <- array(NA,c(h,7))
for (o in 1:omaxTest){
  sweden.trnTest <- head(sweden,22*4-1+o) 
  sweden.tstTest <- tail(sweden,3*4-o+1) 
    for (m in 1:4){
      fit1 <- ets(sweden.trnTest,model=modelsTest[m],damped=dampedTest[m]) 
      frcsTest[,m] <- forecast(fit1,h=h)$mean
      errTest[o,m] <- mean(abs(sweden.tstTest[1:h] - frcsTest[,m]))
    }
  frcsTest[,5] <- tail(sweden.trnTest,frequency(sweden.trnTest))[1:h]
  errTest[o,5] <- mean(abs(sweden.tstTest[1:h] - frcsTest[,5]))
  frcsTest[,6] <- apply(frcsTest[,1:5],1,mean)
  errTest[o,6] <- mean(abs(sweden.tstTest[1:h] - frcsTest[,6]))
  frcsTest[,7] <- apply(frcsTest[,1:5],1,median)
  errTest[o,7] <- mean(abs(sweden.tstTest[1:h] - frcsTest[,7]))
}
colnames(errTest) <- c("AAA","ANA","MNM","MAM","Naive","Comb.Mean","Comb.Median")
errMeanTest <- colMeans(errTest)
errMeanTest
```

```{r}
fit_naive <- naive(sweden.trn)
naive_forecasts <- forecast(fit_naive, h = 4)
mape <- MAPE(sweden.tst,tail(sweden.trnTest,frequency(sweden.trnTest))[1:4])*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Building model with MAM
fit1 <- ets(sweden.trn,model="MNM",damped=FALSE) 
frc1 <- forecast(fit1,h=4)
mape <- MAPE(sweden.tst,frc1$mean)*100
cat("MAPE:", mape, "%\n")
```

```{r}
n <- length(sweden.trn)
X <- array(NA,c(n,5))
for (i in 1:5){
  X[i:n,i] <- sweden.trn[1:(n-i+1)]
}
colnames(X) <- c("y",paste0("lag",1:4))
X <- as.data.frame(X)
plot(X)
fit2 <- lm(y~.,data=X)
fit2 <- step(fit2)
#Forecast on training set
plot(X$y,type="l")
frc2 <- predict(fit2,X)
lines(frc2,col="red")
#Forecast on testing set
frc2 <- array(NA,c(4,1))
for (i in 1:4){
  Xnew <- tail(sweden.trn,4)
  Xnew <- c(Xnew,frc2)
  Xnew <- Xnew[i:(3+i)]
  Xnew <- Xnew[4:1]
  Xnew <- array(Xnew, c(1,4)) 
  colnames(Xnew) <- paste0("lag",1:4)  
  Xnew <- as.data.frame(Xnew)
  frc2[i] <- predict(fit2,Xnew)
} 
mape <- MAPE(sweden.tst,frc2)*100
cat("MAPE:", mape, "%\n")
#MAPE <- mean(abs((sweden.tst - frc2)/sweden.tst)) * 100
#cat("MAPE:", MAPE, "%\n") 
```
```{r}
D <- factor(rep(1:4,24))
X2 <- cbind(X,D)
colnames(X2) <- c(colnames(X2)[1:5],"D")
fit3 <- lm(y~.,data=X2)
summary(fit3)
# Find NA in X2
idx <- is.na(X2)
idx <- rowSums(idx)
idx <- idx == 0
fit_temp <- lm(y~.,data=X2[idx,])
# fit_temp is the same as fit3, without the first NA part 
fit4 <- step(fit_temp)
summary(fit4)
frc3 <- array(NA,c(4,1)) 
for (i in 1:4){
  Xnew <- tail(sweden.trn,4)
  Xnew <- c(Xnew,frc3)
  Xnew <- Xnew[i:(3+i)]
  Xnew <- Xnew[4:1]
  Xnew <- array(Xnew, c(1,4))
  colnames(Xnew) <- paste0("lag",1:4)
  Xnew <- as.data.frame(Xnew)
  D <- as.factor(rep(1:4,1)[i])
  frc3[i] <- predict(fit4,Xnew)
}
mape <- MAPE(sweden.tst,frc3)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Lasso
xx <- as.matrix(X[-(1:4),-1])
# For the target I retain only the first column 
yy <- as.matrix(X[-(1:4),1])
lasso <- cv.glmnet(x=xx,y=yy)
frc4 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(sweden.trn,4),frc4)
  Xnew <- (Xnew[i:(3+i)])[4:1]
  Xnew <- array(Xnew, c(1,4)) 
  colnames(Xnew) <- paste0("lag",1:4) 
  # Forecast
  frc4[i] <- predict(lasso,Xnew)
}
mape <- MAPE(sweden.tst,frc4)*100
cat("MAPE:", mape, "%\n")
```

```{r}
#Ridge
ridge <- cv.glmnet(x=xx,y=yy,alpha=0)
frc5 <- array(NA,c(4,1)) 
for (i in 1:4){
  # Create inputs - note for lasso we do not transform these into data.frame
  Xnew <- c(tail(sweden.trn,4),frc5)
  Xnew <- (Xnew[i:(3+i)])[4:1]
  Xnew <- array(Xnew, c(1,4))
  colnames(Xnew) <- paste0("lag",1:4) # Forecast 
  frc5[i] <- predict(ridge,Xnew)
}
mape <- MAPE(sweden.tst,frc5)*100
cat("MAPE:", mape, "%\n")
```
```{r}
#Building THieF
frc6 <- thief(sweden.trn,h=4,usemodel="arima")
frc6_values <- frc6$mean
mape <- MAPE(sweden.tst,frc6_values)*100
cat("MAPE:", mape, "%\n")
```
```{r}
frc2 <- ts(frc2,frequency=frequency(sweden.tst),start=start(sweden.tst))
frc3 <- ts(frc3,frequency=frequency(sweden.tst),start=start(sweden.tst))
frc4 <- ts(frc4,frequency=frequency(sweden.tst),start=start(sweden.tst))
frc5 <- ts(frc5,frequency=frequency(sweden.tst),start=start(sweden.tst))
frc6 <- ts(frc6_values,frequency=frequency(sweden.tst),start=start(sweden.tst))

# Plot
ts.plot(tail(sweden.trn,4), sweden.tst, fit1$mean, frc2, frc3, frc4, frc5, frc6, col = c("black", "black", "red", "blue","green","yellow","orange","brown"))
legend("bottomleft", c("ETS(MNM)","Stepwise", "Dummies","lasso","Ridge","Thief"), col = c("red", "blue","green", "purple","orange","brown"), lty = 1)
```
